---
title: "White Faced Ibis in the Bear River Migratory Bird Refuge, UT"
format: dashboard
editor: visual
author: "Alastair Buddin"
---

```{r}
#| message: false
library(tidyverse) 
library(janitor)
library(leaflet)
library(leafem)
library(rnaturalearth)
library(sf)
library(terra)
#load and clean data
nest_data_rough <- read.csv("C:/Users/apbb0/OneDrive - Montana State University/STAT408/Project/STAT408_Project_Buddin/wfib.ns.data_data_release.csv")
nest_data = clean_names(nest_data_rough)  


```

```{r}
#| output: false

# #load raster
 #veg_raster = rast("C:/Users/apbb0/OneDrive - Montana State University/STAT408/Project/STAT408_Project_Buddin/LPC_h008v007_HC_2012_V01.tif")
# #reproject raster  
# veg_raster = project(veg_raster, "EPSG:4326")
# #set bounds based on shp
# raster_bounds = vect("C:/Users/apbb0/OneDrive - Montana State University/STAT408/Project/STAT408_Project_Buddin/BRVMBR_bounds.shp")
# raster_bounds = project(raster_bounds, "EPSG:4326")
# #clip raster to shp extent
# veg_raster_clip = crop(veg_raster, raster_bounds, snap="in")
# writeRaster(
#   veg_raster_clip,
#   filename = "veg_raster_clip.tif",
#   overwrite = TRUE,
#   gdal = c("COMPRESS=LZW")
# )


# I had to comment out the above code because the file size was too large

veg_raster_clip <- rast("C:/Users/apbb0/OneDrive - Montana State University/STAT408/Project/STAT408_Project_Buddin/veg_raster_clip.tif")
veg_raster_clip = project(veg_raster_clip, "EPSG:4326" )
# assigns red and nir bands
red_band <- veg_raster_clip[["LPC_h008v007_HC_2012_V01_3"]]
nir_band <- veg_raster_clip[["LPC_h008v007_HC_2012_V01_4"]]
#function to create ndvi raster
calc_ndvi <- function(nir, red){
  ndvi <- (nir-red)/(nir+red)
  return(ndvi)
}
#creating ndvi from clipped raster
ndvi_raster = calc_ndvi(nir_band, red_band)

#making missing values NA
values(ndvi_raster)[is.nan(values(ndvi_raster))] <- NA

ndvi_pal <- colorRampPalette(c(
  "#8c510a",  # brown (bare ground)
  "#d8b365",
  "#f6e8c3",
  "#c7eae5",
  "#5ab4ac",
  "#01665e"   # dark green (dense vegetation)
))

#plot(ndvi_raster, col = ndvi_pal(200))



#saving ndvi to its own .tif
tmpfile <- tempfile(fileext = ".tif")
writeRaster(ndvi_raster, tmpfile, overwrite = TRUE)
#plotting the NDVI  
#plot(ndvi_raster)


#plot(veg_raster_clip)

```

# Overview

Vegetation analysis of nesting habitat for birds can be used to create distribution models and predictiong nesting locations. Comparing an normalized difference vegetation index (NDVI) to vegetation data The nesting data was collected by Ackerman and Herzog in 2020 through the Western Ecological Research Center and USGS, and their analysis was published in the *The Wilson Journal of Ornithology*

## Column {height="80%"}

```{r}
# map wiht ndvi raster
#| fig-width: 14
#| fig-height: 18
map <- leaflet() %>% 
  addTiles()%>%
  addProviderTiles("Esri.WorldImagery")%>%
  addRasterImage(ndvi_raster, colors = ndvi_pal(200))%>%
    setView( lat = 41.474344, lng = -112.296844, zoom = 11)
map
```

# Graphs and Nest Analysis

## Column

```{r}
#basic graph showing Relationship between vegetation height and cover


nest_data_by_veg = nest_data|>
  group_by(veg_percent_vegetation)|>
  mutate(
    survive_count = sum(survive_full)
  )
scale_factor = mean(nest_data$avg_veg_height)/mean(nest_data$survive_count)

ggplot(data= nest_data_by_veg, aes(x = veg_percent_vegetation)) +
geom_point(aes(y = avg_veg_height))+
geom_line(aes(y = survive_count*scale_factor))+
scale_y_continuous(
  name = "Average Vegetation Height (cm)",
  sec.axis = sec_axis(~./scale_factor, name = "Number of Nest Successes")
)




```
